---
title: "Statistical Modeling chap5"
output: pdf_document
---


The differnce in the log likelyhood will be examined. 
$$
\Delta D_{1,2} = -2 \times(log L_1 - log L_2)
$$

The P value will be taken for the confidence interval.

```{r data}
setwd("/Users//shota/git/StatisticalModeling/")
download.file("http://hosho.ees.hokudai.ac.jp/~kubo/stat/iwanamibook/fig/poisson/data3a.csv","data3.csv","curl")
d <- read.csv("data3.csv")
fit2 <- glm(y ~ x, data = d, family = poisson)
fit1 <- glm(formula = y ~ 1, family = poisson, data = d)
```

```{r deviance}
fit2$deviance
fit1$deviance - fit2$deviance
```

The difference in the log likelyhood is about 4.5.

The average number of seeds are estimated as 7.85.
```{r average seeds}
mean(d$y)
fit1
exp(2.058)
```

make 100 models from 7.85

```{r 100 models}
d$y.rnd <- rpois(100, lambda = mean(d$y))
```

Use the glm function to fit the constant model and x model to the new data.

```{r glm new}
fit1 <- glm(y.rnd ~ 1, data = d, family = poisson)
fit2 <- glm(y.rnd ~ x, data = d, family = poisson)
fit1$deviance - fit2$deviance
```

The model with x fits better (smaller deviance) than the actual model (just random poisson).  
Do this 1000 times, and we get a distruibution. This is called the PB method.

```{r pb}
get.dd <- function(d){
  n.sample <- nrow(d)
  y.mean <- mean(d$y)
  d$y.rnd <- rpois(n.sample, lambda = y.mean)
  fit1 <- glm(y.rnd ~ 1, data = d, family = poisson)
  fit2 <- glm(y.rnd ~ 2, data = d, family = poisson)
  fit1$deviance - fit2$deviance
}
pb <- function(d, n.bootstrap){
  replicate(n.bootstrap, get.dd(d))
}
?replicate
getwd()
```

```{r pb method}
source("pb.R")
dd12 <- pb(d, n.bootstrap = 1000)

summary(dd12)
hist(dd12, 100)
sum(dd12 >= 4.5)
```

36 are larger than 4.5.

```{r quantile}
quantile(dd12, 0.95)
```

The 95% confidence interval for this histogram is 3.92.

## Using Chi-Square distruibutions

The distribution of D_1,2 can be approximated by the chi-sq distribution.

```{r chi-sq}
fit1 <- glm(y ~ 1, data = d, family = poisson)
fit2 <- glm(y ~ x, data = d, family = poisson)

anova(fit1, fit2, test = "Chisq")
```

