---
title: "chap2"
output: html_document
---

#「データ解析のための統計モデリング入門」
### Looking into the data
```{r}
setwd("/Users//shota/git/StatisticalModeling/chap2/")
load("data.RData")
data
length(data)
summary(data)
table(data)
par(mfrow=c(1,1))
hist(data, breaks = seq(-0.5, 9.5, 1))
var(data)
sd(data)
sqrt(var(data))
```

### Possion distruibution
The Possion distribution is the distribtuion we want to use.
Lambda is the only parameter in a Possion distribution.
```{r}
y <- 0:9
prob <- dpois(y, lambda = 3.56)
plot(y, prob, type = "b", lty = 2)
cbind(y, prob)
hist(data, breaks = seq(-0.5, 9.5 , 1))
lines(y, 50 * prob, type = "b", lty = 2)
```

Possion distribution
p(y | lambda) = (lambda ^ y) * exp(- lambda) / y!

- The sum is 1: Sigma(y:1 -> Inf) p(y | lambda) = 1
- The mean is lambda
- mean = distribution

Why Possion distribution?
1. Count data
2. Upper limit unknown
3. Mean and distribution are similiar

### Maximum likelihood estimation

Multiplies the possion possiblity of all the subjects of a given lambda.
Estimate the labmda that is best.
L(lambda) = Pi(i: p(y_i | lambda))

The log likelyhood function is thus
log L(lambda) = Sigma i:(yi*log(lambda) - lambda - Sigma k:1 -> yi(log(k)))

In code
```{r}
logL <- function(m) sum(dpois(data, m, log = TRUE))
lambda <- seq(2, 5, 0.1)
plot(lambda, sapply(lambda, logL), type = "l")
```

The maximum lambda can be found by division of log L(lambda) by lambda.

The data was used by rpois()
is this case,
```{r}
rpois(50, 3.5)
```

